The process of retrieving schedule data from the Novasoftware web application

### 2016-04-18 paper/paper.tex
Edits to paper, more README.md

	In the case of the high school I attend; the school provides a separate schedule for Tuesdays via an independent application. This inconvenience of using separate applications for different days of the week further increases the relevance of extracting the Novasoftware data, in order to unify the two schedules into one. Also, when thinking about the implications of having access to this data I realized that it could be used for research purposes; analyzing the schedules of different schools and drawing conclusions from them.

	Any applications are outside the scope of this paper but are mentioned in order to interest the reader and because of their role as motivation for the project.

	\subsection{Purpose}
	%Lite mycket implement här
	The purpose of this paper is provide a complete documentation, as well as a working implementation, of the process of fetching, parsing and extracting data from the Novasoftware public web interface.

	The purpose of the documentation is to act as a guide for implementation of the entirety or parts of the process. An implementation is included, providing a working program capable of extracting current schedule data. This data may for example be used to create custom viewers or for analysis. The documentations also acts as a description of the provided implementation so that it may be examined and reviewed.

	\subsection{Goals}
	The goal of this paper is to provide a documentation of the schedule extraction process that may be implemented in any programming language or platform, as well as including a working implementation, complete with installation and usage instructions.

@@

	The process should be if not in its original form, then with minor adjustments, applicable to every schedule provided by the Novasoftware application. Notes must be included where such adjustment may be neccecary.

	\subsection{Delimitaions}
	Because the Novasoftware schedule data and its formats appears to be at least in part manually input by school representatives with possibly varying conventions, it may not be consistent between schools or classes. Since not every case can be covered, the process described will be adapted to and only guaranteed to work on the current schedules from Värmdö Gymnasium. Despite this; tests on other schedules indicate that the process is applicable to most schedules with none or minor adjustments.

	The process has been developed for the current version of the Novasoftware application, as of the date of publication of this paper, and may not cover future versions.

	For obvious reasons only one implementation is included. The selected language is JavaScript, written for the Node.js platform. The reason behind this decision is convenience; that I feel comfortable with it. JavaScript is also suitable because of its close connection to the web.

@@

	The Novasoftware application is typically accessed via a web browser. A web browser (or simply browser) is a program capable of retrieving and displaying web content. The content is provided by a web server, usually as HTML (Hyper Text Markup Language) to the client (in this case the browser) via HTTP (HyperText Transfer Protocol) over the Internet.\cite{wikipedia web browser} A resource provided by a web server may be called an endpoint. In order to automate the process of fetching resources a different HTTP client than a browser is required. In the implementation Node.js is used to perform the requests. Node.js may be described in short as an environment for running JavaScript outside of the browser.\cite{wikipedia node.js}

	\subsubsection{HTTP-requests}
	An HTTP request is a message sent over the Internet from the client to a server. The server should then return a response containing the requested resource back to the client. HTTP requests are constructed of a URL (Uniform Resource Locator), a method, a set of headers and a request body.

	The URL acts as an address for the requested resource, comparable to how postal addresses are used for mail. A URL and consists of a protocol, a domain, a path and a query string. An example URL is\\
	\url{http://www.novasoftware.se/ImgGen/schedulegenerator.aspx?format=png}

	Here, the protocol is \url{http}, the domain is \url{www.novasoftware.se}, the path is \url{/ImgGen/schedulegenerator.aspx} and finally the query string is \url{?format=png}.
	The protocol denotes the type of request. In this paper only \texttt{http} and its secure counterpart \texttt{https} will be used. The domain specifies the server, while the path points to requested the resource on it. The query string is optional and may contain extra information about the request in a key-value structure. In the example above; the query string tells the server that the \texttt{format} should be \texttt{png}.

	HTTP requests also include a \texttt{method}, where only GET and POST will be used in this paper. The different methods differ in multiple ways. Foremost they include a notion of convention; where GET means that some information is to be fetched and POST means that some data is to be passed to the web server. A POST request may also include data to be passed to the server known as form data because of the traditional application of using forms to send it. In GET requests further information may instead be included in the URL query string, as specified above.

	An HTTP request further includes a set of headers, which provide more information about the request, again in a dictionary (key-value) structure. Headers may be divided into request - and response headers. Request headers sent from the client to the web server, while response headers are sent back to the client in the server response. The request headers may for example provide the server with information about the client, such as the software performing the request, while a response header can indicate for example if the requested action was successful or not. This is often indicated in the \texttt{Status} field. Another relevant header field is the \texttt{Location} response header. If it is provided, along with a 30X (300, 301 etc.) response status, the server is indicating that the resource is instead to be found at the URL provided in value of the \texttt{Location} header.\cite{wikipedia http request}\cite{wikipedia url}

	Another important feature of HTTP requests are that they can include cookies. Cookies are used by applications in order to save small amounts of data in the browser and are sent back and forth in every request. This is often used to give the client an ID (Identifier) so that information about the client can persist between sessions even though HTTP is a stateless protocol.\cite{wikipedia cookie}

	\subsubsection{HTML}
	Because the HTML obtained from the server response is a string, there is a need to transform it into a searchable and more practical structure. Often the document is described in a DOM (Document Object Model), providing a more programmer friendly interface. In order to reference certain locations or data points of the HTML documents, CSS (Cascading Style Sheets) selectors are used, because of their popularity within the web. The method used in this process, where HTML is retrieved and information is extracted from it in an automated program, is often referred to as web scraping.\cite{wikipedia css}\cite{wikipedia web scraping}

	\subsubsection{JSON}
	The output of the program is in the JSON (JavaScript Object Notation) format which is human readable and particularly useful for the web because of it's similarity to JavasScript; the de facto scripting language on the web.\cite{wikipedia json}

>>>>

	In the case of the high school I attend, the school provides a separate schedule for Tuesdays via an independent application. This inconvenience of using separate applications for different days of the week further increases the relevance of extracting the Novasoftware data, in order to unify the two schedules into one. Also, when thinking about the implications of having access to this data I realized that it could be used for research purposes; analyzing the schedules of different schools and drawing conclusions from them.

	Any specific ways to use the retrieved data are outside the scope of this paper; I'm only mentioning these possibilities since they may interest the reader and because of their role as motivation for the project.

	\subsection{Purpose}
	The purpose of this paper is to provide a complete documentation, as well as a working implementation, of the process of fetching, parsing and extracting data from the Novasoftware public web interface.

	The documentation can act a guide for the realization of the whole or parts of the process, while the included implementation provides a working program capable of extracting current schedule data. This data may for example be used to create custom viewers or for analysis. The documentation also acts as a description of the provided implementation so that it may be examined and reviewed.

	\subsection{Goals}
	The goal of the documentation is that it may be implemented in any programming language or platform. My goal with the program is that it should be simple to install and use and that clear instructions will be provided with it.

@@

	The process should be if not in its original form, then with only minor adjustments, appliable to every schedule provided by the Novasoftware application. Notes must be included where such adjustments may be necessary.

	\subsection{Delimitaions}
	Because the Novasoftware schedule data and its formats appears to be at least in part manually inserted by school representatives, possibly with varying conventions, it may not be consistent between schools or classes. Since not every case can be covered here -- the process described will be adapted to and only guaranteed to work -- on the current schedules from Värmdö Gymnasium. Despite this, tests on schedules from other sources indicate that the process is appliable to most schedules with none or only minor adjustments.

	The process has been developed for the current version of the Novasoftware application, at the date of publication of this paper, and may not cover future versions.

	Due to the limitations of this project, only one implementation is included. The selected language is JavaScript, written for the Node.js platform. The reason behind this decision is convenience; the language is widespread and simple. JavaScript is also suitable because of its close connection to the web.

@@

	The Novasoftware application is typically accessed via a web browser (or simply browser), which is a program capable of retrieving and displaying web content. The content is provided by a web server, usually as HTML (Hyper Text Markup Language) to the client (in this case the browser) via HTTP (HyperText Transfer Protocol) over the Internet.\cite{wikipedia web browser} A resource provided by a web server may be called an endpoint. In order to automate the process of fetching resources, a different HTTP client than a browser is required. In the implementation, Node.js is used to perform the requests. Node.js may be described in short as an environment for running JavaScript outside of the browser.\cite{wikipedia node.js}

	\subsubsection{HTTP-requests}
	An HTTP request is a message sent over the Internet from a client to a server; the server should then return a response containing the requested resource back to the client. HTTP requests are constructed of a URL (Uniform Resource Locator), a method, a set of headers and a request body.

	The URL acts as an address for the requested resource, comparable to how postal addresses are used for mail. A URL consists of a protocol, a domain, a path and a query string. An example URL is\\
	\url{http://www.novasoftware.se/ImgGen/schedulegenerator.aspx?format=png}

	Here, the protocol is \url{http}, the domain is \url{www.novasoftware.se}, the path is \url{/ImgGen/schedulegenerator.aspx} and the query string is \url{?format=png}.
	The protocol denotes the type of request. In this paper only \texttt{http} and its secure counterpart \texttt{https} will be used. The domain specifies the server, while the path points to the requested resource on it. The query string is optional and may contain extra information about the request in a key-value structure. In the example above; the query string tells the server that the \texttt{format} should be \texttt{png}.

	HTTP requests also include a \texttt{method}: only GET and POST will be used in this paper. The different methods differ in multiple ways. Primarily, they include a notion of convention, where GET means that some information is to be fetched and POST means that some data is to be passed to the web server. A POST request may also include data to be passed to the server, known as form data because of the traditional application of using forms to send it. In GET requests further information may instead be included in the URL query string, as specified above.

	An HTTP request further includes a set of headers which provide more information about the request, again in a dictionary (key-value) structure. Headers may be divided into request headers and response headers. Request headers are sent from the client to the web server, while response headers are sent back to the client in the server response. The request headers may for example provide the server with information about the client, such as which software is performing the request, while a response header can indicate for example if the requested action was successful or not. This is often indicated in the \texttt{Status} field. Another relevant header field is the \texttt{Location} response header. If this is provided, along with a 30X (300, 301 etc.) response status, the server is indicating that the resource is rather to be found at the URL provided in value of the \texttt{Location} header.\cite{wikipedia http request}\cite{wikipedia url}

	Another important feature of HTTP requests are that they can include cookies. Cookies are used by applications in order to save small amounts of data in the browser, and are sent back and forth in every request. This is often used to give the client an ID (Identifier) so that information about the client can persist between sessions even though HTTP is a stateless protocol.\cite{wikipedia cookie}

	\subsubsection{HTML}
	Because the HTML obtained from the server response is a string, it must be transformed into a searchable and more practical structure. Often the document is described in a DOM (Document Object Model), providing a more programmer friendly interface. In order to reference certain locations or data points of the HTML documents, CSS (Cascading Style Sheets) selectors are used, because of their popularity within the web. The method used in this process, where HTML is retrieved and information is extracted from it in an automated program, is often referred to as web scraping.\cite{wikipedia css}\cite{wikipedia web scraping}

	\subsubsection{JSON}
	The output of the program is in the JSON (JavaScript Object Notation) format which is human readable and particularly useful for the web because of its similarity to JavasScript; the de facto scripting language on the web.\cite{wikipedia json}
